---
title: "gdelt_asd_application"
author: "Franco Caramia"
date: "2023-03-26"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: inline
---

## Introduction

At www.gdeltproject.org, you’ll find the Global Database of Events, Location and Tone. We’d like you to take a look at the data from the week of 9-16 June 2018 and tell us something that the Australian Government might find interesting about it.

```{r, include=F}
rm(list = ls())
gc(full = T)
library(gdeltr2)
library(dplyr)
library(ggplot2)

# Download data (the month of June)

mydate = as.Date("2018-06-30")

dates_to_get = NULL
for (i in 0:30){
  new_date = gsub("-","",mydate - i)
  dates_to_get = c(dates_to_get,  new_date)
}
event_data  = get_data_gdelt_periods_event (periods = dates_to_get)

## Filter out of date events 
event_data = event_data%>% filter(dateEvent>="2018-06-01")



```

## Checking basic event data

**Events Per Day**
```{r}
## Check Events per day

event_data %>% filter(dateEvent>='2018-06-09',dateEvent<="2018-06-16") %>% ggplot(aes(x=dateEvent)) + geom_histogram()
event_data %>% ggplot(aes(x=dateEvent)) + geom_histogram()


```

At first seems like there is a spike in number of events, however, this is just the weekend effect. 

**Check Scores**
```{r}
## Check Events per day

event_data %>% filter(dateEvent>='2018-06-09',dateEvent<="2018-06-16") %>% ggplot(aes(x=scoreGoldstein)) + geom_density() + facet_wrap(~dateEvent)

event_data %>% filter(dateEvent>='2018-06-09',dateEvent<="2018-06-16") %>% ggplot(aes(x=avgTone)) + geom_density() + facet_wrap(~dateEvent)

event_data$dateEventFactor = as.factor(event_data$dateEvent)
event_data %>% filter(dateEvent>='2018-06-09',dateEvent<="2018-06-16") %>% ggplot(aes(x=dateEventFactor,y=scoreGoldstein)) + geom_boxplot()

event_data %>% filter(dateEvent>='2018-06-09',dateEvent<="2018-06-16") %>% ggplot(aes(x=dateEventFactor,y=avgTone)) + geom_boxplot()


```
Seems like looking at total events, there was not much change


**Events related to Australia**


```{r}

event_data %>% 
  filter(idCountryActor1=='AS'|idCountryActor2=='AS'|idCountryAction=='AS')  %>%
  filter(dateEvent>='2018-06-09',dateEvent<="2018-06-16") %>% ggplot(aes(x=dateEvent)) + geom_histogram()

event_data %>% 
  filter(idCountryActor1=='AS'|idCountryActor2=='AS'|idCountryAction=='AS')  %>%
  ggplot(aes(x=dateEvent)) + geom_histogram()


event_data %>%
  filter(idCountryActor1=='AS'|idCountryActor2=='AS'|idCountryAction=='AS') %>%
  filter(dateEvent>='2018-06-09',dateEvent<="2018-06-16") %>% 
  ggplot(aes(x=scoreGoldstein)) + geom_density() + facet_wrap(~dateEvent)

event_data %>%
  filter(idCountryActor1=='AS'|idCountryActor2=='AS'|idCountryAction=='AS') %>%
  filter(dateEvent>='2018-06-09',dateEvent<="2018-06-16") %>% 
  ggplot(aes(x=avgTone)) + geom_density() + facet_wrap(~dateEvent)


event_data %>% 
  filter(idCountryActor1=='AS'|idCountryActor2=='AS'|idCountryAction=='AS') %>%
  filter(dateEvent>='2018-06-09',dateEvent<="2018-06-16") %>% 
  ggplot(aes(x=dateEventFactor,y=scoreGoldstein)) + geom_boxplot()

event_data %>% 
  filter(idCountryActor1=='AS'|idCountryActor2=='AS'|idCountryAction=='AS') %>%
  filter(dateEvent>='2018-06-09',dateEvent<="2018-06-16") %>% 
  ggplot(aes(x=dateEventFactor,y=avgTone)) + geom_boxplot()


event_data %>%
  filter(idCountryActor1=='AS'|idCountryActor2=='AS'|idCountryAction=='AS') %>%
  filter(dateEvent>='2018-06-09',dateEvent<="2018-06-16") %>% 
  ggplot(aes(x=countMentions)) + geom_density() + facet_wrap(~dateEvent)



```

Checking the distributions in the data. 

**Break down by event description**

```{r}

event_data %>%
  filter(idCountryActor1=='AS'|idCountryActor2=='AS'|idCountryAction=='AS') %>%
  filter(dateEvent>='2018-06-09',dateEvent<="2018-06-16") %>% 
  ggplot(aes(y=scoreGoldstein,x=dateEventFactor)) + geom_boxplot() + facet_wrap(~nameQuad)

event_data %>%
  filter(idCountryActor1=='AS'|idCountryActor2=='AS'|idCountryAction=='AS') %>%
  filter(dateEvent>='2018-06-09',dateEvent<="2018-06-16") %>% 
  ggplot(aes(y=avgTone,x=dateEventFactor)) + geom_boxplot() + facet_wrap(~nameQuad)



```
Cooperations are usually positive and conflicts usually negative in both Goldstein and avgTone 


**Event type frequencies**

```{r}

event_data %>% 
  filter(dateEvent>='2018-06-09',dateEvent<="2018-06-16") %>% 
  ggplot(aes(x=dateEvent)) + geom_histogram() + facet_wrap(~nameQuad)

event_data %>% 
  ggplot(aes(x=dateEvent)) + geom_histogram()+ facet_wrap(~nameQuad)


event_data %>% 
  filter(idCountryActor1=='AS'|idCountryActor2=='AS'|idCountryAction=='AS')  %>%
  filter(dateEvent>='2018-06-09',dateEvent<="2018-06-16") %>% 
  ggplot(aes(x=dateEvent)) + geom_histogram() + facet_wrap(~nameQuad)

event_data %>% 
  filter(idCountryActor1=='AS'|idCountryActor2=='AS'|idCountryAction=='AS')  %>%
  ggplot(aes(x=dateEvent)) + geom_histogram()+ facet_wrap(~nameQuad)


```


**Are metrics correlated? **

```{r}

library(corrplot)
library(tidyverse)

corr_mat = event_data %>% select(avgTone,scoreGoldstein,countMentions,countSources,countArticles) 
corrplot.mixed(cor(corr_mat,use="complete.obs"))

```

We have learned so far that the number of events and the distribution of scores (Goldstein,AvgTone) doesnt change much. 

avgTone and Goldstein are weakly correlated, so it might be better to separate them. 

Mentions, Sources and Articles are highly correlated, so we can use only one

Next:

Can we find events that have mix of "out of ordinary" (-2>zscores>2) in all of Goldstein, avgTone, mentions/sources/articles?

The hypothesis is that if an event has very high/low impact score and high/low frequency, it should be important. 

Let's do this for the world and also for events related to Australia. 

**Calculate Zscores and search for events**
```{r}

tone_important_events = event_data %>%
  filter(dateEvent>='2018-06-09',dateEvent<="2018-06-16") %>%
  mutate(z_score = (avgTone-mean(avgTone,na.rm = T))/sd(avgTone,na.rm=T)) %>%
  filter(z_score>=2|z_score<=-2) %>% 
  pull(idGlobalEvent)


goldstein_important_events = event_data %>%
  filter(dateEvent>='2018-06-09',dateEvent<="2018-06-16") %>%
  mutate(z_score = (scoreGoldstein-mean(scoreGoldstein,na.rm = T))/sd(scoreGoldstein,na.rm=T)) %>%
  filter(z_score>=2|z_score<=2) %>%
  pull(idGlobalEvent)

## Only highly mentioned events
mentions_important_events = event_data %>%
  filter(dateEvent>='2018-06-09',dateEvent<="2018-06-16") %>%
  mutate(z_score = (countMentions-mean(countMentions,na.rm = T))/sd(countMentions,na.rm=T)) %>%
  filter(z_score>=2) %>%
  pull(idGlobalEvent)


very_important = intersect(tone_important_events,goldstein_important_events)
very_important = intersect(very_important,mentions_important_events)

# 518 events are either highly positive/negative and high number of mentions 

# Any mention Australia? 

important_event_data = event_data %>% filter(idGlobalEvent%in%very_important)

important_event_data_au = important_event_data %>%
 filter(idCountryActor1=='AS'|idCountryActor2=='AS'|idCountryAction=='AS')  

# 6 events are related to Australia

```


Found 518 events with a highly positive or negative score and high following (mentions)
6 them possibly related to Australia 
However, some links are not working and one article is unrelated to Australia due to a victim named "Sydney".

I concluded that the addition of the Jama’at Mujahideen Bangladesh terrorist group by the Department of Home Affairs 
to the Australian terrorist group list might be of great interest to the ADF. 

